---
title: "PML project"
author: "Patrick Ball"
date: "Thursday, April 16, 2015"
output: html_document
---

### Load packages and the dataset

```{r}
library(dplyr); library(ggplot2)
library(caret); library(randomForest)
if (getwd() != "C:/Users/Patrick/Coursera/pml_project") {
    setwd("C:/Users/Patrick/Coursera/pml_project")
}
train <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")
```

### Clean up the Data
The sparse nature of many columns and the size of the data are problematic (given my processing capabilities). The following steps resolve these issues:

- Remove time & user information from data
- Remove columns primarily populated with NA
- Remove columns primarily populated with "" entries

```{r}
train <- select(train, -(X:num_window))
train <- train[,colSums(is.na(train))==0]
train <- train[,colSums(train[]=="")==0]

test <- select(test, -(X:num_window))
test <- test[,colSums(is.na(test))==0]
test <- test[,colSums(test[]=="")==0]
```

### Fit a CART Model

I want to fit at least two models. The CART model is a decent place to start for predicting a non-binary factor response.

```{r, cache=TRUE}
cartModel <- train(classe~., method="rpart", data=train)
cartPred <- predict(cartModel, train[,-53])
table(cartPred, train$classe)
```

The table of predictions vs. reality doesn't look very good. There are lots of entries off of the diagonal. The model information below includes the accuracy:

```{r}
cartModel
```

An accuracy rate of 0.51 is very low.

### Fit a Random Forest Model

The random forest offers higher accuracy, but may overfit.

```{r, cache=TRUE}
rfModel <- randomForest(train[,-53], train[,53])
rfPred <- predict(rfModel, train[,-53])
table(rfPred, train$classe)
```

This produced perfect prediction of the training data.

```{r}
rfModel
```

The out-of-bag error estimate is 0.32%. That's very strong. Hopefully this hasn't overfitted the data.

### Create Test Data Predictions

First, create and look at the predictions using the random forest model.

```{r}
rfTest <- predict(rfModel, test[,-53])
table(rfTest)
```

Notice that these results are skewed toward A and B results. That may be problematic. Let's check the CART predictions.

```{r}
cartTest <- predict(cartModel, test[,-53])
table(cartTest)
```

These results are even more skewed that the predictions from the random forest model, and look less likely, so stick with the random forest model.

### Create Files

```{r}
answers <- as.character(rfTest)
write_files <- function(x) {
    for(i in 1:length(x)) {
        filename <- paste0("problem_id_",i,".txt")
        write.table(x[i], file=filename, quote=FALSE,
                    row.names=FALSE, col.names=FALSE)
    }
}
write_files(answers)
```